---
title: 'Stats 250: Helpful Functions'
author: "Instructional Team"
date: "Week of April 10th"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Helpful Functions

First, load in the stats250package and the penguins data set! 

```{r stats250package}
library(stats250package)
```

```{r readPenguins}
penguins <- read.csv("penguins.csv")
```


Then, check out the various procedures and corresponding functions below.


### Difference in Two Means

We may wish to test if flipper length differs, on average, between Adelie and Gentoo penguins. 

This is a little bit tricky, because our data set contains three species, but we can subset the data to include our two species of interest using the following code:

```{r subsetToTwoGroups}
penguins_subset <- subset(penguins, species == "Adelie" | species == "Gentoo")
```

- Be careful with spelling and capitalization
- Note the use of the double equals sign (==)
- A line (|) is included to filter to both species (Adelie or Gentoo)


We could then use this subset (that contains the two species of interest) to descriptively compare the flipper lengths.

```{r boxplot}
boxplot(penguins_subset$flipper_length_mm ~ penguins_subset$species,
        xlab = "Species",
        ylab = "Flipper Length (in mm)",
        main = "Boxplot of Flipper Length by Species")
```

- Variables of interest (quantitative by categorical)
- Labels and title


We then need to subset the data to check the assumptions.

```{r subsetAdelie}
adelie <- subset(penguins, species == "Adelie")
```

```{r subsetGentoo}
gentoo <- subset(penguins, species == "Gentoo")
```

```{r QQPlotAdelie}
qqnorm(adelie$flipper_length_mm, main = "QQ Plot of Flipper Lengths (Adelie)")
qqline(adelie$flipper_length_mm)
```

```{r QQPlotGentoo}
qqnorm(gentoo$flipper_length_mm, main = "QQ Plot of Flipper Lengths (Gentoo)")
qqline(gentoo$flipper_length_mm)
```


And then finally, we could run the test for a difference in two means. 

```{r twoMeansHT}
t.test(penguins_subset$flipper_length_mm ~ penguins_subset$species,
       mu = 0,
       alternative = "two.sided")
```

- Variables of interest (quantitative by categorical)
- Hypothesized difference
- Direction of the alternative hypothesis ("less", "greater", or "two.sided")


***


### Difference in Two Proportions

We may wish to test if the population proportion of male penguins is different between Adelie and Gentoo penguins.

We could subset the data to include the two groups of interest (as done in the first code chunk for a difference in means), but it isn't necessary.

```{r twowayTable}
addmargins(table(penguins$species, penguins$sex))
```

Using this table we can calculate the two sample proportions and also calculate the estimate of the common population proportion. For our assumptions, we could use R as a calculator to make sure we have at least 10 success and failures for each species under the null hypothesis.

Then, we would run the test using the `twoPopPropHT()` function. While we haven't used this yet, it is very similar to the `onePopPropHT()` function from Lab 4. 

```{r twoPopPropHT}
twoPopPropHT(x1 = 80, n1 = 100, x2 = 180, n2 = 200, alt = "two.sided")
```

- x1 = Number of successes (Group 1)
- n1 = Sample size (Group 1)
- x2 = Number of successes (Group 2)
- n2 = Sample size (Group 2)
- Direction of the alternative hypothesis ("less", "greater", or "two.sided")


***


### Linear Relationship

We may wish to test if there is a linear relationship between flipper length and body mass. 

We would start with a scatterplot.

```{r scatterplot}
plot(body_mass_g ~ flipper_length_mm,
     data = penguins,
     xlab = "Flipper Length (in mm)",
     ylab = "Body Mass (in g)",
     main = "Scatterplot of Body Mass vs Flipper Length")
```

- Response variable by explanatory variable
- Data set of interest


In order to check the assumptions for regression, we would actually have to start by creating our model first!

```{r linearModel}
lm_penguins <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

- Response variable by explanatory variable
- Data set of interest


Then we can use the plot function and the stored linear model to create our two plots used to help check our assumptions. 

```{r residualPlot}
plot(lm_penguins, which = 1)
```

```{r qqPlotofResiduals}
plot(lm_penguins, which = 2)
```

- which = 1 gives us the residual plot
- which = 2 gives us the qq plot of the residuals


Finally, we can use the summary() function to run the test:

```{r regressionTest}
summary(lm_penguins)
```


